A Domain-Driven, Generative Data Model for BigPetStore

### 2014-07-23 bdcloud_paper/latex/paper.tex
edits

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Domain-Driven, Generative Data Model for BigPetStore}

@@

\begin{abstract}
Both developers and users of big data systems benefit from realistic example applications.  For developers, such applications are useful for testing, benchmarking, and evaluating design choices; for users, example applications provide starting points for learning methods, developing their own applications, and implementing new analytics workflows.

@@

%Access to high-quality, semantically-rich data sets and complex, realistic analytics workflow examples that can scale from local development environments to large clusters are key for consistent development, testing, and benchmarking of big data software, algorithms, and infrastructure. BigPetStore is a big data application blueprint, which incorporates a data generator and complex examples of cleaning, aggregation, and analytics, built around a fictional chain of pet store which has been used for testing and benchmarking in the Apache BigTop distribution. We present a new generative data model for BigPetStore and demonstrate the model's ability to generate semantically-rich data that can be used for complex analytics workflows at variable scale ranging from a single machine to a large cluster.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

% no keywords

@@

\section{Introduction}
Big data applications process large, continuously-changing data sets with the general goal of information and knowledge extraction.  With the wide variety of big data tools available and lagging documentation, both developers and users of big data systems benefit from the availability of realistic example applications.  For developers, such applications are useful for testing, benchmarking, and evaluating design choices; for users, example applications provide starting points for learning methods, developing their own applications, and implementing new analytics workflows.

BigPetStore is a big data application blueprint built around processing transaction data for a fictional chain of pet stores.  BigPetStore mainly targets the Hadoop \cite{Hadoop} ecosystem with examples implemented for loading, cleaning, aggregating, and performing analytics on data using Hive \cite{Thusoo2010}, Pig \cite{Olston2008,Gates2009}, and Mahout \cite{Mahout}. BigPetStore has been incorporated into the open-source Apache BigTop distribution \cite{BigTop}, where it is used for testing and as a reference application. One of BigPetStore's key features is the ability to scale from a single machine to a large cluster, making it easy to develop projects on a local machine and transfer the application to a large cluster for production testing.

@@

A variety of approaches for generating synthetic data set exist.  TeraGen and the Intel Hadoop Benchmark Suite \cite{Huang2010} are popular tools that can generate data sets quickly and at any scale, but the resulting data is semantically-void and not useful for much more than simple benchmarks.  Multiple frameworks exist for generating synthetic data sets that satisfy relational database schemas \cite{Ghazal2013,Rabl2011a,Frank2012,Rabl2011,Gray1994,Bruno2005,Hoag2007}. Several approaches even provide domain-specific languages \cite{Bruno2005,Hoag2007} for specifying additional constraints such as which distributions to sample from and allow for modeling basic relationships between fields using simple arithmetic equations \cite{Alexandrov2012}. Arasu, et al. \cite{Arasu2011} demonstrated that constraint-solving techniques could be used as an alternative to procedural approaches. These frameworks allow for reproducing the structural properties of real data, but the frameworks are not expressive enough to describe the dynamic generation processes and latent variables that would be needed to embed the desired informational complexity and rich semantics needed for BigPetStore's analytics examples.

@@

Until such methods have reached maturity, BigPetStore's needs are best met by a customized, domain-driven model. Another example of such a model is The Information Discovery and Analysis System (IDAS) Data Set Generator (IDSG) \cite{Jeske2005,Lin2006}. IDSG was developed to to test the effectiveness of IDASs in identifying potential terrorism threat scenarios using synthetic data.  Like BigPetStore, the developers of IDSG needed to avoid licensing restrictions and privacy issues associated with real data.

@@

In this work, we present a new domain-driven model and simulation method for BigPetStore.  Compared with BigPetStore's previous data generator, our model and implementation can generate data which contains geospatial, temporal, and quantitative features, similar to the type of data which businesses and organizations might typically encounter.  Combining the features of the TeraGen (scalability)  and MovieLens \cite{MovieLens} (semantically rich) input data sets commonly used for big data benchmarking, we present a synthetic data set generator which can be used to benchmark lower level tasks (such as sorting) as well as higher level tasks (such as clustering, recommending, and search indexing) at arbitrary scale.  To demonstrate that the model generates data with desirable properties, we performed an example analysis designed to guide a fictional advertising campaign (a typical business use case). The model's implementation available as open source software.

@@

\subsection{Description of Generated Data}
The model generates data for four types of records: stores, customers, transactions, and transaction items (Figure~\ref{fig:relational-data-model}).  Store records contain both a unique identifier and a location in the form of a 5-digit zip code. Customer records consist of a unique identifier, a name, and a location given as a 5-digit zip code. Transaction records consist of a customer identifier, a store identifier, and a time given in days since the beginning of the simulation. Transaction item records contain a transaction identifier and an item description.  The item description is a list of key-value pairs stored as a string.  Key-value pairs are used since item characteristics differ depending on the type of item.  For example, dog food has a brand, a flavor, a size (in lbs), and a per-unit cost, while poop bags have a brand, a color, a count, and a per-unit cost.

@@

\begin{align*}
p(\text{location}=z) = &Z^-1 p(\text{location}=z | \text{population}(z)) \\
&p(\text{location}=z | \text{income}(z))
\end{align*}

@@

The current implementation was prototyped in Python.  We are working to implement the model using Hadoop and Spark.  We will commit the resulting implementations to the Apache BigTop project to enable ease of access and immediate benefit to current users.

>>>>

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{A Domain-Driven, Generative Data Model for BigPetStore}

@@

\begin{abstract}
Generating large amounts of semantically-rich data for testing big data workflows is paramount for scalable performance benchmarking and quality assurance in modern machine-learning and analytics workloads.  The most obvious use case for such a generative algorithm is in conjunction with a big data application blueprint, which can be used by developers (to test their emerging big data solutions) as well as end users (as a starting point for validating infrastructure installations, building novel applications, and learning analytics methods).

%Both developers and users of big data systems benefit from realistic example applications.  For developers, such applications are useful for testing, benchmarking, and evaluating design choices; for users, example applications provide starting points for learning methods, developing their own applications, and implementing new analytics workflows.

@@

\end{abstract}

@@

\section{Introduction}
Big data applications process large, dynamics, multidimensional data sets with the general goal of information and knowledge extraction.  With the wide variety of big data tools available and lagging documentation, both developers and users of big data systems benefit from the availability of realistic example applications.  For developers, such applications are useful for testing, benchmarking, and evaluating design choices; for users, example applications provide starting points for learning methods, developing their own applications, and implementing new analytics workflows.

BigPetStore is one such big data application blueprint built around processing transaction data for a fictional chain of pet stores.  BigPetStore targets the Hadoop \cite{Hadoop} ecosystem with examples implemented for loading, cleaning, aggregating, and performing analytics on data using Hive \cite{Thusoo2010}, Pig \cite{Olston2008,Gates2009}, and Mahout \cite{Mahout}. BigPetStore has been incorporated into the open-source Apache BigTop distribution \cite{BigTop}, where it is used for testing and as a reference application. One of BigPetStore's key features is the ability to scale from a single machine to a large cluster, making it easy to develop projects on a local machine and transfer the application to a large cluster for production testing.

@@

A variety of approaches for generating synthetic data sets exist.  TeraGen and the Intel Hadoop Benchmark Suite \cite{Huang2010} are popular tools that can generate data sets quickly and at any scale, but the resulting data is semantically-void and not useful for much more than simple benchmarks.  Multiple frameworks exist for generating synthetic data sets that satisfy relational database schemas \cite{Ghazal2013,Rabl2011a,Frank2012,Rabl2011,Gray1994,Bruno2005,Hoag2007}, and several approaches even provide domain-specific languages \cite{Bruno2005,Hoag2007} for specifying additional constraints (such as which distributions to sample from and allow for modeling basic relationships between fields using simple arithmetic equations \cite{Alexandrov2012}).

Recent work has addressed some aspects of the need for more dynamic data set generation: Arasu, et al. \cite{Arasu2011} demonstrated that constraint-solving techniques could be used as an alternative to procedural approaches. Such frameworks allow for reproducing the structural properties of real data, but the frameworks are not expressive enough to describe the dynamic generation processes and latent variables that would be needed to embed the desired informational complexity and rich semantics needed for BigPetStore's analytics examples.

@@

Until such purely generic methods reach maturity, BigPetStore's needs are best met by a customized, domain-driven model. Another example of such a model is The Information Discovery and Analysis System (IDAS) Data Set Generator (IDSG) \cite{Jeske2005,Lin2006}. IDSG was developed to to test the effectiveness of IDASs in identifying potential terrorism threat scenarios using synthetic data.  Like BigPetStore, the developers of IDSG needed to avoid licensing restrictions and privacy issues associated with real data.

@@

In this work, we present a new domain-driven model and simulation method for BigPetStore.  Compared with BigPetStore's previous data generator, our model and implementation can generate data which contains geospatial, temporal, and quantitative features, similar to the type of data which businesses and organizations might typically encounter.  Combining the features of the TeraGen (scalability)  and MovieLens \cite{MovieLens} (semantically rich) input data sets commonly used for big data benchmarking, we present a synthetic data set generator which can be used to benchmark lower level tasks (such as sorting) as well as higher level tasks (such as clustering, recommending, and search indexing) at arbitrary scale.  To demonstrate that the model generates data with desirable properties, we performed an example analysis designed to guide a fictional advertising campaign (a typical business use case). The model's implementation was made available as open source software.

@@

\subsection{Description of Generated Data}
The model generates data for four types of records: stores, customers, transactions, and transaction items (Figure~\ref{fig:relational-data-model}).  Store records contain both a unique identifier and a location in the form of a 5-digit zip code. Customer records consist of a unique identifier, a name, and a location given as a 5-digit zip code. Transaction records consist of a transaction identifier, a customer identifier, a store identifier, and a time given in days since the beginning of the simulation. Transaction item records contain a transaction identifier and an item description.  The item description is a list of key-value pairs stored as a string.  Key-value pairs are used since item characteristics differ depending on the type of item.  For example, dog food has a brand, a flavor, a size (in lbs), and a per-unit cost, while poop bags have a brand, a color, a count, and a per-unit cost.

@@

\begin{align*}
p(\text{location}=z) = &Z p(\text{location}=z | \text{population}(z)) \\
&p(\text{location}=z | \text{income}(z))
\end{align*}

@@

The current implementation was prototyped in Python.  We are working to implement the model using Hadoop and Spark.  We will commit the resulting implementations to BigPetStore hosted in the Apache BigTop distribution to enable ease of access and immediate benefit to current users.
